---
title: Model Results for GDP ~ Energy
author: Grant Esparza
date: "4/26/2018"
---

```{r}

## Libraries used
library(lme4)
library(sjPlot)
library(dplyr)
library(ggplot2)

```

```{r}

#Read data
states.energy <- read.csv("~/Energy_Analysis/data/usa_states_energy.csv")
usa.energy <- read.csv("~/Energy_Analysis/data/usa_energy.csv")

## Clean it up a little
states.energy <- states.energy %>% select(c(year, State, CLTCB, FFTCB, GETCB, HYTCB,
                                          MGTCB, NGTCB, SOTCB, WYTCB, GDP,
                                          GDP.total))
names(states.energy)[11] <- "GDP.state"
states.energy$GDP.total.growth <- ifelse(usa.energy[38:56, 22] > 1.03,
                                         1,0)

names(states.energy)

```

**Some plotting showed that in order to keep these variables on the same scale
as GDP, I needed to use a log transformation. Since all variables are transformed
the analysis should be valid.**

```{r, fig.width=5, fig.height=5, fig.align='center'}

## Perform a log transformation on all variables
for (i in 1:8){
    states.energy[, i+2] <- log(states.energy[,i+2] + 1)
}
states.energy$GDP.total <- log(states.energy$GDP.total)
states.energy$GDP.state <- log(states.energy$GDP.state)

```

**After constant tweaking I'm pretty much in the same place in regards to the 
random intercept model. I talked to Edward about my model a little bit and he
was also at a loss as to why the variance for the random intercept would be 
equal to zero. This happens when including the intercept (_see no '-1'_).

However when removing the intercept, we get meaningful results that seem to 
reflect what I've been able to generate in plots(_see the density plots at the
end_.) Edward says this could be a failing of frequentist statistics, and that 
for now we should run with _log(GDP.state) ~ -1 + log(HYTCB) + (1|State)_. We
should ask Robin about this tomorrow.**


```{r, fig.width=5, fig.height=5, fig.align='center'}

## Partial Pooling
contin.model.intercept <- lmer(GDP.state ~ HYTCB + (1 | State),
                                data=states.energy)


contin.model.no_intercept <- lmer(GDP.state ~ -1 + HYTCB +
                                  (1 | State), data=states.energy)

summary(contin.model.intercept)

## Make sjPlot for random effects
plot_model(contin.model.intercept, sort.est="(Intercept)", type="re",
           y.offset=.4)

plot_model(contin.model.no_intercept, sort.est="(Intercept)", type="re", 
           y.offset=.4)

summary(contin.model.no_intercept)

contin.model1 <- lmer(GDP.state ~ -1 + CLTCB + FFTCB + MGTCB + NGTCB+
                      GETCB + HYTCB + SOTCB + WYTCB + (1 | State), 
                     data=states.energy)
plot_model(contin.model1, sort.est="(Intercept)", type="re", 
           y.offset=.4)

summary(contin.model1)

states.energy$renewable <- rowSums(states.energy %>% 
                                   select(HYTCB, SOTCB, WYTCB, GETCB))
states.energy$non_renewable <- rowSums(states.energy %>% 
                                   select(CLTCB, FFTCB, MGTCB, NGTCB))


states.energy$non_renewable <- log(states.energy$non_renewable)
states.energy$renewable <- log(states.energy$renewable)

contin.model.type <- lmer(GDP.state ~ -1 + renewable + non_renewable + (1 | State), 
                     data=states.energy)

plot_model(contin.model.type, sort.est="(Intercept)", type="re", 
           y.offset=.4)


summary(contin.model.type)

```

**As you can see in the plots below, HYTCB seems to be a prevelant contributor
compared to other states. This reflects what the random intercept model shows us
in the previous sjPlot. Another apparent feature of these graphs is the overwhelming
domination of nonrewnewable sources...but we already could have guessed that 
coming in.**

```{r, fig.width=7, fig.height=7, fig.align='center'}

ggplot(states.energy) +
  geom_density(aes(x=CLTCB), fill="black", alpha=.5) +
  geom_density(aes(x=FFTCB), fill="brown", alpha=.5) + 
  geom_density(aes(x=GETCB), fill="green", alpha=.5) +
  geom_density(aes(x=HYTCB), fill="blue", alpha=.5) +
  geom_density(aes(x=MGTCB), fill="gray", alpha=.5) +
  geom_density(aes(x=NGTCB), fill="yellow", alpha=.5) +
  geom_density(aes(x=SOTCB), fill="orange", alpha=.5) +
  geom_density(aes(x=WYTCB), fill="light blue", alpha=.5) +
  facet_wrap(~State)

```

**I also went ahead and ran a factor analysis to see if nonrewnewable and 
renewable energy ended up in distinct groups or "factors". At least in regards
to their total consumption the plot below shows some pretty distinct grouping.**

```{r, fig.width=5, fig.height=5, fig.align='center'}

# Factor Analysis
library(corrplot)
library(psych)

states.dat <- states.energy %>% select(c(CLTCB, FFTCB, GETCB, HYTCB, MGTCB,
                                         NGTCB, SOTCB, WYTCB))

extract_pca <- princomp(states.dat, cor=TRUE)
var_pc <- (extract_pca$sdev)^2
qplot(x=1:length(var_pc), y=var_pc, geom=c("point", "line")) +
    xlab("PC number") + ylab("Eigenvalue")

pc.model.oblimin <- principal(states.dat, nfactors=2, rotate="oblimin",
                              scores=TRUE)
print(pc.model.oblimin)

load <- pc.model.oblimin$loadings
plot(load, type="n")
text(load, labels=rownames(load))

```

**All in all, if we can run with the second variation of the random intercept 
model we can come up with some decent results. If Robin says we can't use it 
I still have the tables from the for loops running lm on each energy source but
I realllllly don't like how those turned out. 
